---
title: "Graduation rates report"
author: "Feifan Cao"
date: "12/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#all the packages we used in this project
library(ggplot2)
library(dplyr)
```

## Introduction
In this project, we are interested in if graduation rate related to different district and different year in Iowa. We got our data from https://data.iowa.gov/Primary-Secondary-Ed/4-Year-Graduation-Rates-in-Iowa-by-Cohort-and-Publ/tqti-3w6t.This dataset provides the 4-Year graduation rates in Iowa by cohort (represented by graduating class) and public school district starting with the Class of 2009.
```{r}
gr=read.csv("4-Year_Graduation_Rates_in_Iowa_by_Cohort_and_Public_School_District.csv")
```


## Data  
```{r}
dim(gr)
variable.names(gr)
```

There are 3566 rows with 8 columns in our data set. Here are the introductions for each variable:

Graduating Class - Year the class is expected to graduate

Fall Freshman Year - The year the cohort entered their freshman year of high school

District - ID given by the Department of Education for the school district

District Name - Name of the school district.

Graduates - First time freshman in year indicated in fall freshman year and transfers into the cohort in grades 9 to 12 who graduated in year indicated in graduating class or earlier. 

Total Cohort - First time freshman in year indicated in fall freshman year and transfers into the cohort in grades 9 to 12 less those who transferred out (including deceased). 

Graduation Rate - Graduates divided by Total Cohort multiplied by 100.

## Graphs
We would like to use Graduation Rate as our dependent variable and we would like to analysis different district and freshman year based on the graphs. 

In the first graph, we are interested in if there is relationship between district and graduation rate. We use group_by function to summarize the the variable `District`, and then we use summarize and mutate function to find the mean graduation rate of year 2009-2019. In the end, we draw a bar plot for of different District and mean Graduation rate.
```{r}
gr %>% 
  group_by(District) %>% 
  summarize(Graduation.Rate= mean(Graduation.Rate)) %>%
  mutate(District= reorder(District, -Graduation.Rate, mean)) %>%
  ggplot(aes(District, weight=Graduation.Rate)) + geom_bar()
```

```{r}

```

