---
title: "Graduation rates report"
author: "Feifan Cao, Chuxiao Yu"
date: "12/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#all the packages we used in this project
library(ggplot2)
library(dplyr)
library(lubridate)
library(patchwork) 
library(hrbrthemes)
```

## Introduction
In this project, we are interested in if graduation rate related to different district and different year in Iowa. We got our data from https://data.iowa.gov/Primary-Secondary-Ed/4-Year-Graduation-Rates-in-Iowa-by-Cohort-and-Publ/tqti-3w6t.This dataset provides the 4-Year graduation rates in Iowa by cohort (represented by graduating class) and public school district starting with the Class of 2009.
```{r}
df=read.csv("4-Year_Graduation_Rates_in_Iowa_by_Cohort_and_Public_School_District.csv")
```


## Data  
```{r}
dim(df)
variable.names(df)
length(unique(df$District)) 
length(unique(df$District.Name))
range(df$Graduating.Class)

```

There are 3566 rows with 8 columns in our data set. Here are the introductions for each variable:

Graduating Class - Year the class is expected to graduate. We can see that the graduation year is from 2009 to 2019.

Fall Freshman Year - The year the cohort entered their freshman year of high school

District - ID given by the Department of Education for the school district. There are 348 different district ID in total.

District Name - Name of the school district. There are  378 different district names in total. There is more District Name(`District.Name`) than the district ID(`District`) we assume there are some district names share the same district ID or maybe the district name have been changed during 2009-2019 but using the same District ID.

Graduates - First time freshman in year indicated in fall freshman year and transfers into the cohort in grades 9 to 12 who graduated in year indicated in graduating class or earlier. 

Total Cohort - First time freshman in year indicated in fall freshman year and transfers into the cohort in grades 9 to 12 less those who transferred out (including deceased). 

Graduation Rate - Graduates divided by Total Cohort multiplied by 100.

## Clean the data set
```{r}
gr=na.omit(df)
head(gr)
dim(gr)
```
Since there are a lot of NA values, we decide to delete all the rows with NA. After cleaning, there are 2458 rows left in our data set. The new data set is called gr in R code.

## Graphs
We would like to use Graduation Rate as our dependent variable and we would like to analysis different district and freshman year based on the graphs. 

In the first graph, we are interested in if there is relationship between district and graduation rate. We use group_by function to summarize the the variable `District`, and then we use summarize and mutate function to find the mean graduation rate of year 2009-2019. In the end, we draw a bar plot for of different District and mean Graduation rate.

Based on the output graph, it is too concentrated to tell if there is any relationship between district and graduation rate.
```{r}
gr %>% 
  group_by(District) %>% 
  summarize(Graduation.Rate= mean(Graduation.Rate)) %>%
  mutate(District= reorder(District, -Graduation.Rate, mean)) %>%
  ggplot(aes(District, weight=Graduation.Rate)) + geom_bar()
```

In the following graph, we would like to see if there is any relationship between the year of graduate (`Graduating.Class`) and graduation rate(`Graduation.Rate`). First of all, we need to  change the gradating.class into character variable to move on. We used group_by function first to divide the data set into according year and use ggplot to draw the relationship between graduation year and graduation rate in a boxplot to analysis. 
Based on the output box plot, we can see that graduation class do have a little effect on the graduation but not significant. We can see that the median graduation and 25% quantile is not affecting by graduation year but the 75% quantile has a increasing trend from graduation class 2011-2016. We can also see that there is always a few outliers each year with low graduating rate.
```{r}
gr$Graduating.Class=as.character(gr$Graduating.Class)
gr %>% 
  group_by(Graduating.Class) %>% 
  ggplot(aes(Graduating.Class,y=Graduation.Rate))+geom_boxplot()
```

## Ames
We are interested in the graduation rate and total cohort in Ames and we created the following graphs. And we can see that 2018 has the highest graduation rate and 2010 has the lowest graduation rate. We can also conclude that for Ames, there is no interaction between Graduating class, total cohort or graduation rate.
```{r}
ames=gr%>%filter(District.Name=="Ames")%>%
  group_by(Graduating.Class)

ames%>%ggplot(aes(x=Graduating.Class, y=Total.Cohort, fill=Graduation.Rate))+geom_bar(stat="identity")

ggplot(ames, aes(x=Graduating.Class)) +
 geom_bar( aes(y=Total.Cohort), stat="identity", size=.1, fill="plum", alpha=.4)+ 
  geom_point(aes(y=(Graduation.Rate-91)*50), size=2) +
  scale_y_continuous(
    name = "Total.Cohort",
    sec.axis = sec_axis( trans=~./50+91, name="Graduation.Rate")
  ) +theme_ipsum()



```

## Related Data Set 2

### Intro and data cleaning
After all the researches before, we would like to learn further, thus in this related data set (graduate intention), we are going to learn about where did the students go after they graduate from high school. We got our data set from https://educateiowa.gov/document-type/graduate-intentions-district-including-graduate-counts. Since there is no data for year 2009, 2010, 2011, and 2019, we are going to only focus on year 2012-2018. There were multiple excel forms and we have already combined them together and added a new column called `Graduation.Year`, in which year that the students graduates.

In the first part of the code, we are introducing the data, change all the N/A and empty cell into NA, and delete all the NA values for further research. We also change the data type to what they are supposed to be. The final data set is saved as gi.After cleaning the data, we can see that there are 1859 rows with 15 columns.
```{r}
df3 <- read.csv("graduate_intention.csv")
#change all the N/A and empty cell into NA and delete them
df3[df3 == ""] <- NA
df3[df3 == "N/A"] <- NA
gi=na.omit(df3)
variable.names(gi)
str(gi)
gi$Diploma.Count=as.numeric(gi$Diploma.Count)
gi$Private.4.Year.College=as.numeric(gi$Private.4.Year.College)
gi$Public.4.Year.College=as.numeric(gi$Public.4.Year.College)
gi$Private.2.Year.College=as.numeric(gi$Private.2.Year.College)
gi$Community.College=as.numeric(gi$Community.College)
gi$Other.Training=as.numeric(gi$Other.Training)
gi$Employment=as.numeric(gi$Employment)
gi$Homemaker=as.numeric(gi$Homemaker)
gi$Military=as.numeric(gi$Military)
gi$Unknown=as.numeric(gi$Unknown)
gi$Graduation.Year=as.character(gi$Graduation.Year)
str(gi)
gi=na.omit(gi)
dim(gi)
```

### basic analysis
Next, we would like to some basic analysis for this data. 

According to the output, we can see that most students will go to college(private 4 year college, public 4 year college, private 2 year college and community college) after they graduate. Employment is another popular choice for students who graduate from high school.
```{r}
gi%>%group_by(Graduation.Year)%>%
  summarize(
    sum.Diploma.Count=sum(Diploma.Count),
    sum.Private.4.Year.College=sum(Private.4.Year.College),
    sum.Public.4.Year.College=sum(Public.4.Year.College),
    sum.Private.2.Year.College=sum(Private.2.Year.College),
    sum.Community.College=sum(Community.College),
    sum.Other.Training=sum(Other.Training),
    sum.Employment=sum(Employment),
    sum.Homemaker=sum(Homemaker),
    sum.Military=sum(Military),
    sum.Unknown=sum(Unknown)
  )
```
According to the graph graduation.year versus diploma count below, we can see that graduation is not a factor that affects diploma count.
```{r}
ggplot(gi,aes(x=Graduation.Year, y=Diploma.Count))+geom_boxplot()
```

```{r}
ames2=gi%>%filter(District.Name=="Ames")%>%group_by(Graduation.Year)
#2012
slices12=c(64, 150, 63, 3, 1, 12, 1, 2, 18)
lbls12=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices12, labels = lbls12, main="Pie Chart of Countries")

#2013
slices13=c(49, 154, 76, 2, 7, 16, 1, 4, 0)
lbls13=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices13, labels = lbls13, main="Pie Chart of Countries")
```

