---
title: "DS 202 Final Project"
output: 
  github_document:
    number_section: FALSE
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, warning = TRUE)


knit_hooks$set( fig.cap = function(before, options, envir) {
  if ( before ) {
    chr <- sprintf("<p><small><strong><a name='fig:%s' value = %d>Figure %d</a></strong>: %s</small></p>", options$label, options$fig.num, options$fig.num, options$fig.cap)
    chr <- sprintf("<p><small><strong><a name='fig:%s'>%s</a></strong>: %s</small></p>", options$label, options$label,  options$fig.cap)
    
    return(chr)
  }
})


chunkref <- local({
  function(chunklabel) {
    sprintf('[%s](#%s)', chunklabel, paste0("fig:",chunklabel) )
  }  
})

library(ggplot2)
library(dplyr)
library(lubridate)
library(patchwork) 
library(hrbrthemes)
```


This repository serves as a starter repo for your final project, and this Rmd is supposed to serve as a starter file for your project report.

## Part I: Repo Structure {.unnumbered}

The structure sketched out below is an idea of what your repository might look like. You can use it as a starting base and change according to your needs. But think about the changes that you make!

    -- code
    |   |   -- any R scripts you need but don't want to include directly in the write-up
    -- data
    |   |   -- csv files (cleaned data)
    -- data-raw
    |   |   -- raw data files 
    |   |   -- data description files, origin
    |   |   -- Codebook
    -- final-project.Rmd
    -- images  # only images that are not created by the Rmd
    -- LICENSE
    -- README.md
    -- README.Rmd
    -- README_files # folder with files created during the knitting process

## Part II: Project report {.unnumbered}



# Title of your project {.unnumbered}

Authors: Chuxiao Yu, Feifan Cao, Haoran Wang, Wenzheng Wang

## Abstract (TL;DR) {.unnumbered}

An abstract is a quick summary of your work. Ideally it should motivate someone to read the rest of the paper. Include one sentence each on

-   what is the project about?
-   what is the motivation for doing it?
-   what data is your work based on? and where does it come from? = what are your main findings? (one sentence each)

# Intro/Background/Motivation

What is the topic of your project, why is it relevant?

At the end of the Intro, write a sentence describing what each of the (result) sections is about, e.g. in section [Results 1] we show the relationship between XXX and YYY,  section [Results 2] also considers the effect of variable ZZZ. ...
Finally we conclude with a quick summary of our findings and potential follow-up work in section [Conclusions].

Somewhere at the beginning of your project, include a code chunk that includes all of the R packages you are using throughout. In this document, the setup code chunk is called `setup` (see line 8) Also make sure to set defaults for the code chunks - like should they be visible? (probably not: echo=FALSE). Do you want to automatically include warnings? (probably yes, for creating the Rmd, to make sure that all warnings are accounted for)

Our project is about researching the graduation rates related to different graduation years, different district, and different genders. The graduation rate not only gives people insight into how many students are finishing their degrees in a timely manner once they enroll, it is also a transparent metric that holds a school accountable and helps meassure the quality of the school.

# Quick Data Summary

```{r}
#graduation rate data
df=read.csv("4-Year_Graduation_Rates_in_Iowa_by_Cohort_and_Public_School_District.csv")
dim(df)
variable.names(df)
length(unique(df$District)) 
length(unique(df$District.Name))
range(df$Graduating.Class)
gr=na.omit(df)
head(gr)
dim(gr)
#graduation intention data
df3 <- read.csv("graduate_intention.csv")
df3[df3 == ""] <- NA
df3[df3 == "N/A"] <- NA
gi=na.omit(df3)
variable.names(gi)
str(gi)
gi$County=as.character(gi$County)
gi$District=as.character(gi$District)
gi$Diploma.Count=as.numeric(gi$Diploma.Count)
gi$Private.4.Year.College=as.numeric(gi$Private.4.Year.College)
gi$Public.4.Year.College=as.numeric(gi$Public.4.Year.College)
gi$Private.2.Year.College=as.numeric(gi$Private.2.Year.College)
gi$Community.College=as.numeric(gi$Community.College)
gi$Other.Training=as.numeric(gi$Other.Training)
gi$Employment=as.numeric(gi$Employment)
gi$Homemaker=as.numeric(gi$Homemaker)
gi$Military=as.numeric(gi$Military)
gi$Unknown=as.numeric(gi$Unknown)
gi$Graduation.Year=as.character(gi$Graduation.Year)
str(gi)
gi=na.omit(gi)
dim(gi)
```


# Results

Each line of exploration is supposed to be featured in one of the Results sections. Make sure to change to more interesting section headers!

<<<<<<< HEAD
## Results 1-results after analyzing graduation rate data

```{r}
#graduation rate vs district
gr %>% 
  group_by(District) %>% 
  summarize(Graduation.Rate= mean(Graduation.Rate)) %>%
  mutate(District= reorder(District, -Graduation.Rate, mean)) %>%
  ggplot(aes(District, weight=Graduation.Rate)) + geom_bar()
#graduation rate summary by district name
gr%>%
  group_by(District.Name)%>%
  summarize(avg.gr=mean(gr$Graduation.Rate))
#graduation rate vs graduating class
gr$Graduating.Class=as.character(gr$Graduating.Class)
gr %>% 
  group_by(Graduating.Class) %>% 
  ggplot(aes(Graduating.Class,y=Graduation.Rate))+geom_boxplot()
#ames graduation rate
ames=gr%>%filter(District.Name=="Ames")%>%
  group_by(Graduating.Class)

ames%>%ggplot(aes(x=Graduating.Class, y=Total.Cohort, fill=Graduation.Rate))+geom_bar(stat="identity")

ggplot(ames, aes(x=Graduating.Class)) +
 geom_bar( aes(y=Total.Cohort), stat="identity", size=.1, fill="plum", alpha=.4)+ 
  geom_point(aes(y=(Graduation.Rate-91)*50), size=2) +
  scale_y_continuous(
    name = "Total.Cohort",
    sec.axis = sec_axis( trans=~./50+91, name="Graduation.Rate")
  ) +theme_ipsum()


```



## Results 2

## Results 3-results after analyzing graduation intention

```{r}
#summarize by graduation year
gi%>%group_by(Graduation.Year)%>%
  summarize(
    sum.Diploma.Count=sum(Diploma.Count),
    sum.Private.4.Year.College=sum(Private.4.Year.College),
    sum.Public.4.Year.College=sum(Public.4.Year.College),
    sum.Private.2.Year.College=sum(Private.2.Year.College),
    sum.Community.College=sum(Community.College),
    sum.Other.Training=sum(Other.Training),
    sum.Employment=sum(Employment),
    sum.Homemaker=sum(Homemaker),
    sum.Military=sum(Military),
    sum.Unknown=sum(Unknown)
  )
#graduation year vs diploma count
ggplot(gi,aes(x=Graduation.Year, y=Diploma.Count))+geom_boxplot()
#county vs diploma count
gi %>%
  group_by(County) %>%
  ggplot(aes(x=County,y=Diploma.Count, color=Graduation.Year))+geom_bar(stat="identity")
gi %>%
  group_by(County) %>%
  summarize(
    sum.diploma=sum(Diploma.Count))
    
#pie charts
ames2=gi%>%filter(District.Name=="Ames")%>%group_by(Graduation.Year)
#2012
slices12=c(64, 150, 63, 3, 1, 12, 1, 2, 18)
lbls12=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices12, labels = lbls12, main="Pie Chart of Graduates Intention")

#2013
slices13=c(49, 154, 76, 2, 7, 16, 1, 4, 0)
lbls13=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices13, labels = lbls13, main="Pie Chart of Graduates Intention")

#2014
slices14=c(54,157,54,4,3,20,0,8,0)
lbls14=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices14, labels = lbls14, main="Pie Chart of Graduates Intention")

#2015
slices15=c(36,159,48,1,5,29,2,0,1)
lbls15=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices15, labels = lbls15, main="Pie Chart of Graduates Intention")

#2017
slices17=c(35,137,79,0,5,32,0,4,13)
lbls17=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices17, labels = lbls17, main="Pie Chart of Graduates Intention")

```


# Conclusions

Give a quick summary of your work. Here is the place to be a bit critical and discuss potential limitations. Add a sentence on what else you would have liked to include in your data exploration if you had more time or more members in your team. 

## Data source {.unnumbered}

Where does the data come from, who owns the data? Where are all the scripts that you need to clean the data?

## References {.unnumbered}

List all resources you used.
