---
title: "DS 202 Final Project"
output: 
  github_document:
    number_section: FALSE
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, warning = TRUE)


knit_hooks$set( fig.cap = function(before, options, envir) {
  if ( before ) {
    chr <- sprintf("<p><small><strong><a name='fig:%s' value = %d>Figure %d</a></strong>: %s</small></p>", options$label, options$fig.num, options$fig.num, options$fig.cap)
    chr <- sprintf("<p><small><strong><a name='fig:%s'>%s</a></strong>: %s</small></p>", options$label, options$label,  options$fig.cap)
    
    return(chr)
  }
})


chunkref <- local({
  function(chunklabel) {
    sprintf('[%s](#%s)', chunklabel, paste0("fig:",chunklabel) )
  }  
})

library(ggplot2)
library(dplyr)
library(lubridate)
library(patchwork) 
library(hrbrthemes)
```


This repository serves as a starter repo for your final project, and this Rmd is supposed to serve as a starter file for your project report.

## Part I: Repo Structure {.unnumbered}

The structure sketched out below is an idea of what your repository might look like. You can use it as a starting base and change according to your needs. But think about the changes that you make!

    -- code
    |   |   -- any R scripts you need but don't want to include directly in the write-up
    -- data
    |   |   -- csv files (cleaned data)
    -- data-raw
    |   |   -- raw data files 
    |   |   -- data description files, origin
    |   |   -- Codebook
    -- final-project.Rmd
    -- images  # only images that are not created by the Rmd
    -- LICENSE
    -- README.md
    -- README.Rmd
    -- README_files # folder with files created during the knitting process

## Part II: Project report {.unnumbered}



# Iowa High School Graduation Rate {.unnumbered}

Authors: Chuxiao Yu, Feifan Cao, Haoran Wang, Wenzheng Wang

## Abstract (TL;DR) {.unnumbered}

In this project, we are interested in if graduation rates related to different graduation years, different district, and different genders, and what are graduates intention. The graduation rate gives students insight into how many students are finishing their degrees in a timely manner once they enroll. It not only is a transparent metric that holds a school accountable, but it also can help a student measure the quality of the school. 

We have three different data sets. The first one is about graduation rate. It is from data.iowa.gov. We are trying to find relationship between graduation rate and district, graduation year. The second data is about dropout rate. It is from educateiowa.gov. We are trying to find the dropout rate trend. The third data is about graduation intention. It is from educateiowa.gov. We are trying to find the graduates intention proportion.


# Intro/Background/Motivation

What is the topic of your project, why is it relevant?

At the end of the Intro, write a sentence describing what each of the (result) sections is about, e.g. in section [Results 1] we show the relationship between XXX and YYY,  section [Results 2] also considers the effect of variable ZZZ. ...
Finally we conclude with a quick summary of our findings and potential follow-up work in section [Conclusions].

Somewhere at the beginning of your project, include a code chunk that includes all of the R packages you are using throughout. In this document, the setup code chunk is called `setup` (see line 8) Also make sure to set defaults for the code chunks - like should they be visible? (probably not: echo=FALSE). Do you want to automatically include warnings? (probably yes, for creating the Rmd, to make sure that all warnings are accounted for)

Our project is about researching the graduation rates related to different graduation years, different district, and different genders. The graduation rate not only gives people insight into how many students are finishing their degrees in a timely manner once they enroll, it is also a transparent metric that holds a school accountable and helps meassure the quality of the school.

In section 'result 1-results after analyzing graduation rate data', we show the relationship between graduation rate and district and graduating class. Section2 'result 2', we show the trend of dropout rate and the relationship between dropout rate and gender. In section 'Results 3-results after analyzing graduation intention', we show the relationship between diploma count and county and graduation year, and a the trend of graduates intention includes a pie charts (for Ames) about proportion of graduates intention.

# Quick Data Summary

```{r, include=FALSE}
#graduation rate data
df=read.csv("4-Year_Graduation_Rates_in_Iowa_by_Cohort_and_Public_School_District.csv")
dim(df)
variable.names(df)
length(unique(df$District)) 
length(unique(df$District.Name))
range(df$Graduating.Class)
gr=na.omit(df)
head(gr)
dim(gr)
#graduation intention data
df3 <- read.csv("graduate_intention.csv")
df3[df3 == ""] <- NA
df3[df3 == "N/A"] <- NA
gi=na.omit(df3)
variable.names(gi)
str(gi)
gi$County=as.character(gi$County)
gi$District=as.character(gi$District)
gi$Diploma.Count=as.numeric(gi$Diploma.Count)
gi$Private.4.Year.College=as.numeric(gi$Private.4.Year.College)
gi$Public.4.Year.College=as.numeric(gi$Public.4.Year.College)
gi$Private.2.Year.College=as.numeric(gi$Private.2.Year.College)
gi$Community.College=as.numeric(gi$Community.College)
gi$Other.Training=as.numeric(gi$Other.Training)
gi$Employment=as.numeric(gi$Employment)
gi$Homemaker=as.numeric(gi$Homemaker)
gi$Military=as.numeric(gi$Military)
gi$Unknown=as.numeric(gi$Unknown)
gi$Graduation.Year=as.character(gi$Graduation.Year)
str(gi)
gi=na.omit(gi)
dim(gi)
```

```{r, include=FALSE}
#dropout data
library(tidyverse)
library(readxl)

# dropout analysis
# load and clean data
# 1. read xlsx file
d2020<-read_excel("DR712_2021_District.xlsx",skip=3)
d2019<-read_excel("dr712bydistrict2020.xlsx",skip=3)
d2018<-read_excel("dr712bydistrict2019.xlsx",skip=3)
d2017<-read_excel("Dropout Rates by district.xlsx",skip=3)
d2016<-read_excel("2016-2017 Grade 7-12 Dropout Rates by District and Subgroup updated 4-12-18.xlsx",skip=7)

# 2. remove useless columns
d2016[45:45] <- list(NULL)
d2016[9:38] <- list(NULL)
d2017[9:50] <- list(NULL)
d2018[9:38] <- list(NULL)
d2019[9:38] <- list(NULL)
d2020[9:38] <- list(NULL)
d2016[1:3] <- list(NULL)
d2017[1:3] <- list(NULL)
d2018[1:3] <- list(NULL)
d2019[1:3] <- list(NULL)
d2020[1:3] <- list(NULL)

# 3. remove useless rows
d2016 <- d2016[-(334:335),]
d2017 <- d2017[-319,]
d2018 <- d2018[-316,]
d2019 <- d2019[-(315:316),]
d2020 <- d2020[-(314:315),]

# change column name
colnames(d2016)<-c("district","district_name","overall_dropout","overall_enrollments","overall_rate","f_dropout","f_enrollments","f_rate","m_dropout","m_enrollments","m_rate")
colnames(d2017)<-c("district","district_name","overall_dropout","overall_enrollments","overall_rate","f_dropout","f_enrollments","f_rate","m_dropout","m_enrollments","m_rate")
colnames(d2018)<-c("district","district_name","overall_dropout","overall_enrollments","overall_rate","f_dropout","f_enrollments","f_rate","m_dropout","m_enrollments","m_rate")
colnames(d2019)<-c("district","district_name","overall_dropout","overall_enrollments","overall_rate","f_dropout","f_enrollments","f_rate","m_dropout","m_enrollments","m_rate")
colnames(d2020)<-c("district","district_name","overall_dropout","overall_enrollments","overall_rate","f_dropout","f_enrollments","f_rate","m_dropout","m_enrollments","m_rate")

# 4. add year to each dataset
d2016 <- d2016 %>% mutate(year=2016)
d2017 <- d2017 %>% mutate(year=2017)
d2018 <- d2018 %>% mutate(year=2018)
d2019 <- d2019 %>% mutate(year=2019)
d2020 <- d2020 %>% mutate(year=2020)

# 5. rbind datasets 
df <- rbind(d2016,d2017)
df <- rbind(df,d2018)
df <- rbind(df,d2019)
df <- rbind(df,d2020)

# 6. drop rows with **** in overall rate
df <-df[!(df$overall_rate=="****"),]

# 7. convert variables type
df <- transform(df, district= as.numeric(district),
                overall_dropout=as.numeric(overall_dropout),
                overall_enrollments= as.numeric(overall_enrollments),
                overall_rate= as.numeric(overall_rate),
                f_dropout= as.numeric(f_dropout),
                f_enrollments= as.numeric(f_enrollments),
                f_rate= as.numeric(f_rate),
                m_dropout= as.numeric(m_dropout),
                m_enrollments= as.numeric(m_enrollments),
                m_rate= as.numeric(m_rate))

# 8. replace NA with 0
df <- df %>% replace(is.na(.), 0)

# 9. create df with split gender
female <- df %>% select(district, f_dropout,f_enrollments,f_rate, year)
female <- female %>% mutate(gender="female")
colnames(female)<-c("district", "dropout" , "enrollments" , "rate" , "year", "gender")
male <- df %>% select(district, m_dropout,m_enrollments,m_rate, year)
male <- male %>% mutate(gender="male")
colnames(male)<-c("district", "dropout" , "enrollments" , "rate" , "year", "gender")
df_gender <- rbind(female, male)

# 10. create dataframe with accumulated overall value
df2 <- (df %>% group_by(year) %>%
          mutate(dropout=sum(overall_dropout),
                 enrollment=sum(overall_enrollments), 
                 year = year)) %>% 
  select(year, dropout, enrollment)

df2 <- df2[!duplicated(df2), ] %>% mutate(rate_percentage = (dropout/enrollment))
```

# Results

Each line of exploration is supposed to be featured in one of the Results sections. Make sure to change to more interesting section headers!

## Results 1-results after analyzing graduation rate data

```{r}
#graduation rate vs district
gr %>% 
  group_by(District) %>% 
  summarize(Graduation.Rate= mean(Graduation.Rate)) %>%
  mutate(District= reorder(District, -Graduation.Rate, mean)) %>%
  ggplot(aes(District, weight=Graduation.Rate)) + geom_bar()
#graduation rate summary by district name
gr%>%
  group_by(District.Name)%>%
  summarize(avg.gr=mean(gr$Graduation.Rate))
#graduation rate vs graduating class
gr$Graduating.Class=as.character(gr$Graduating.Class)
gr %>% 
  group_by(Graduating.Class) %>% 
  ggplot(aes(Graduating.Class,y=Graduation.Rate))+geom_boxplot()
#ames graduation rate
ames=gr%>%filter(District.Name=="Ames")%>%
  group_by(Graduating.Class)

ames%>%ggplot(aes(x=Graduating.Class, y=Total.Cohort, fill=Graduation.Rate))+geom_bar(stat="identity")

ggplot(ames, aes(x=Graduating.Class)) +
 geom_bar( aes(y=Total.Cohort), stat="identity", size=.1, fill="plum", alpha=.4)+ 
  geom_point(aes(y=(Graduation.Rate-91)*50), size=2) +
  scale_y_continuous(
    name = "Total.Cohort",
    sec.axis = sec_axis( trans=~./50+91, name="Graduation.Rate")
  ) +theme_ipsum()


```


For graduation rate vs district: based on the output graph, it is too concentrated to tell if there is any relationship between district and graduation rate.

For graduation rate vs graduating class: Graduation class do have a little effect on the graduation but not significant. The median graduation and 75% quantile is not affecting by graduation year but the 25% quantile has a increasing trend from graduation class 2011-2016.

For Ames graduation rate: both total cohort and graduation class are not factors that affects graduation rate


## Results 2
```{r}
# 1. overall dropout by year
df_gender %>% group_by(year) %>% ggplot(aes(x=year, y=dropout, fill=gender)) + geom_bar(stat = "identity")

# 2. overal enrollments by year
df_gender %>% group_by(year) %>% ggplot(aes(x=year, y=enrollments, fill=gender)) + geom_bar(stat = "identity")
```
These two bar plot shows overall number of enrollments and dropout by year. 
We notice that the overall enrollments is increating by year, but overall dropout is decreasing from 2016 to 2019, and it suddenly increase a lot in 2020. This is pretty strange, there should be some big event happened there. We guess that it is because the appearance of covid-19. 
On the other hand, the ratio of female and male enrollments are close to 50 50, however, as for dropouts, male obviously account for a greater proportion. It means that male has higher dropout rate than female in IOWA.

```{r}
df2 %>% ggplot(aes(x=year,y=rate_percentage))+geom_point()+geom_line()+ylim(0.01, 0.021)
```
The next plot shows total dropout rate by year, which is a changing form of the previous plot. It display a more obvious trend. 
```{r, include=FALSE}
options(digits = 5) 
df3 <- data.frame()
for (i in 2:5){
  df3 <-rbind(df3, c(df2$year[i], as.numeric(sprintf((df2$rate_percentage[i] - df2$rate_percentage[i-1]) , fmt = '%#.5f'))))
}
colnames(df3)<-c("year", "rate_change")
```
```{r}
# 3. dropout rate changed by year
df3 %>% ggplot(aes(x=year, y=rate_change)) + geom_bar(stat = "identity") + geom_text(aes(label=rate_change))
```
In order to view the rate of change by year, we made this plot.This is the accurate change of overall dropout rate by year. Same as what we see in the previous plot, the rate change are negative  from 2016 to 2019, and a big increasing appear in 2020.

```{r}
# 4.  overall dropout rate by district shape
df_gender %>% group_by(district) %>% 
  ggplot(aes(x=reorder(district, rate) , 
             y=rate/nrow(df), fill=gender)) + 
  geom_bar(stat = "identity")
```
This plot shows ranked mean dropout rate by district colored by gender. The mean rate is calculated by accumulated overall rate divided by number of districts which is the numer of rows. we can see the distribution of dropout rate by district. 
```{r}
# 5. overall dropout number by district shape
df_gender %>% group_by(district) %>% 
  ggplot(aes(x=reorder(district, dropout) , 
             y=dropout, fill=gender)) + 
  geom_bar(stat = "identity")
```
Comparing with next plot which shows dropout count by district, we can see that the distribution is pretty different.The gap between districts is larger. Most of the districts have very low dropouts, there are only a new extreme large value. One district even has about 3000 overall dropout. It is very abnormal to have such an outlier, but it is hard to find out what caused the out lier based on the current data.

## Results 3-contribution to overall dropout based on number of dropout region
```{r, include=FALSE}
count_values <- function(df, lower, upper, year) {
  sum(df$overall_dropout >= lower & df$overall_dropout <= upper & df$year == year)
}

df6 <- data.frame()
for (i in 2016:2020){
  df6 <- rbind(df6, c(i, count_values(df,0,10, i)))
  df6 <- rbind(df6, c(i,count_values(df,10,20, i)))
  df6 <- rbind(df6, c(i,count_values(df,20,50, i)))
  df6 <- rbind(df6, c(i,count_values(df,50,100, i)))
  df6 <- rbind(df6, c(i,count_values(df,100,200, i)))
  df6 <- rbind(df6, c(i,count_values(df,200,500, i)))
  df6 <- rbind(df6, c(i,count_values(df,500,10000, i)))
}
colnames(df6) <- c("year", "count")
c("year", "0~10","10~20","20~50","50~100","100~200","200~500","500+")
```
```{r}
# 6. number of district of different number of overall dropout students
pie((df6%>%filter(year==2016))$count, labels = c("0~10", "10~20", "20~50", "50~100", "100~200", "200~500", "500+"), main="number of dropout(2016)")
pie((df6%>%filter(year==2017))$count, labels = c("0~10", "10~20", "20~50", "50~100", "100~200", "200~500", "500+"), main="number of dropout(2017)")
pie((df6%>%filter(year==2018))$count, labels = c("0~10", "10~20", "20~50", "50~100", "100~200", "200~500", "500+"), main="number of dropout(2018)")
pie((df6%>%filter(year==2019))$count, labels = c("0~10", "10~20", "20~50", "50~100", "100~200", "200~500", "500+"), main="number of dropout(2019)")
pie((df6%>%filter(year==2020))$count, labels = c("0~10", "10~20", "20~50", "50~100", "100~200", "200~500", "500+"), main="number of dropout(2020)")
```
Next, we divided number of dropouts into different regions  as listed below, and made pie plot by year. We can see that district with less than 10 dropouts each year actually takes more than 75 percent of the total dropout. But in 2020, the percentage suddenly decreased. and number of dropout between 10 and 20 obviously increased, and number of dropout larger than 20 also increased a little bit. this means instead of only several districts, the increase in number of dopout in 2020 is is a large-scale phenomenon.


## Results 4-results after analyzing graduation intention

```{r}
#summarize by graduation year
gi%>%group_by(Graduation.Year)%>%
  summarize(
    sum.Diploma.Count=sum(Diploma.Count),
    sum.Private.4.Year.College=sum(Private.4.Year.College),
    sum.Public.4.Year.College=sum(Public.4.Year.College),
    sum.Private.2.Year.College=sum(Private.2.Year.College),
    sum.Community.College=sum(Community.College),
    sum.Other.Training=sum(Other.Training),
    sum.Employment=sum(Employment),
    sum.Homemaker=sum(Homemaker),
    sum.Military=sum(Military),
    sum.Unknown=sum(Unknown)
  )
#graduation year vs diploma count
ggplot(gi,aes(x=Graduation.Year, y=Diploma.Count))+geom_boxplot()
#county vs diploma count
gi %>%
  group_by(County) %>%
  ggplot(aes(x=County,y=Diploma.Count, color=Graduation.Year))+geom_bar(stat="identity")
gi %>%
  group_by(County) %>%
  summarize(
    sum.diploma=sum(Diploma.Count))
    
#pie charts
ames2=gi%>%filter(District.Name=="Ames")%>%group_by(Graduation.Year)
#2012
slices12=c(64, 150, 63, 3, 1, 12, 1, 2, 18)
lbls12=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices12, labels = lbls12, main="Pie Chart of Graduates Intention")

#2013
slices13=c(49, 154, 76, 2, 7, 16, 1, 4, 0)
lbls13=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices13, labels = lbls13, main="Pie Chart of Graduates Intention")

#2014
slices14=c(54,157,54,4,3,20,0,8,0)
lbls14=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices14, labels = lbls14, main="Pie Chart of Graduates Intention")

#2015
slices15=c(36,159,48,1,5,29,2,0,1)
lbls15=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices15, labels = lbls15, main="Pie Chart of Graduates Intention")

#2017
slices17=c(35,137,79,0,5,32,0,4,13)
lbls17=c("Private.4.Year.College", "Public.4.Year.College", "Private.2.Year.College", "Community.College", "Other.Training","Employment", "Homemaker",  "Military", "Unknown" )
pie(slices17, labels = lbls17, main="Pie Chart of Graduates Intention")

```


Summarize by graduation year: Most students will go to college(Private.4.Year.College, Public.4.Year.College, Private.2.Year.College, Community.College are considered college) after graduate. 
Employment is a popular choice for those who are not going to college.

Graduation year vs diploma count: Graduation year is not a factor that affects diploma count

County vs diploma count: County is a factor that affect diploma count. County 77 has the most diploma in year 2012-2018.This does not mean County 77 has the highest graduation rate because we don’t know how many `Total.Cohort` (as it called in the graduation data) in this county

Pie charts: most graduates would like to go to college.4 year public college is the most popular choice, about half of graduates will go to 4 year public college every year. 


# Conclusions

Give a quick summary of your work. Here is the place to be a bit critical and discuss potential limitations. Add a sentence on what else you would have liked to include in your data exploration if you had more time or more members in your team. 

From our analysis, we find that graduation class do have a little effect on the graduation but not significant. 
For Ames, both total cohort and graduation class are not factors that affects graduation rate. 
The overall dropout are decreasing from 2016 to 2019, but suddenly increased in 2020, while the overall enrollments is increasing. We also notice that The district with number of dropouts below than 10 each year contributes more than 75% of total dropouts.However, we find some connection with the districts and dropouts but not able to seek a very sighnificant finding on this part.


If we had more time and more data we could make analysis of the dropout rate with covid data



## Data source {.unnumbered}

Where does the data come from, who owns the data? Where are all the scripts that you need to clean the data?

The Iowa government website owns the data.

https://data.iowa.gov/Primary-Secondary-Ed/4-Year-Graduation-Rates-in-Iowa-by-Cohort-and-Publ/tqti-3w6t.This

https://educateiowa.gov/document-type/graduate-intentions-district-including-graduate-counts

https://educateiowa.gov/graduation-rates-and-dropout-rates

We get these 3 datasets from Iowa government website and all the scripts can be found in 'Quick Data Summary' above.

## References {.unnumbered}

List all resources you used.

https://www.earnest.com/blog/graduation-and-retention-rates/
